Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-05-23 doc/bcspaper/article.tex
sorry, adding the subcaption and soul packages back in
Maria methods edits

%\usepackage{subcaption}
%\captionsetup{compatibility=false}
%\usepackage{soul}

@@

\section{Methods}
\label{sec:2}

%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting)

The first process in the processing pipeline is conversion of the papers to a
format that can be analysed by SAPIENTA.  Currently, SAPIENTA supports the
SciXML and PubMed Central DTDs.  Papers in PDF format must be converted to XML
in order to be processed.  For this, we use PDFX \cite{PDFX}, a free service
hosted by the University of Manchester (\url{http://pdfx.cs.man.ac.uk/}).  PDFX
also provides output that already has some metadata, such as title, author, and
abstract, associated with it.  SAPIENTA allocates each sentence within a paper
a separate CoreSC label.  However, the sentence boundaries must be detected
before the document is passed into SAPIENTA for annotation.  The sentence
splitter uses a regular expression to split the sentences within the document
and then adds the necessary \verb|<s>| tags to the markup to indicate the
location of each sentence.  Once the paper has been split, SAPIENTA is used to
annotate each sentence with a relevant CoreSC label.  These labels are
calculated based upon the sentence's location within the paper as well pairs
and triplets of words found consecutively within the sentence.

%how types were chosen,

Open Access (OA) journals are research journals that provide their papers free
to read online, often making revenue by charging authors to publish their
articles \cite{solomon2012study}.  Many OA journals provide free read access to
their papers, but do not permit the user of articles for data mining purposes.
The PLoS journals\footnote{PLoS: \url{http://www.plos.org/}} contains large
volumes of OA literature under a permissive license that allows data mining.
PLoS use the Pubmed Central markup schema, which is compatible with SAPIENTA,
for papers published through their journal.  The PLoS journals advanced search
offers approximately 50 types of paper through which to restrict the search.
Many of these paper type categories contain too few papers to be useful. Others
are too specific (e.g.``Message from PLoS''). We chose to look at XX paper
types (Essay, Correspondence, Synopsis, Perspective, Viewpoint, Opinion,
Review, Research). These paper types represent a range of types, some of which
we expect to overlap or have an unclear distinction.

%how papers were acquired 
The papers were downloaded from PLoS using {\em plosget.py}, a short python
script that uses the PLoS restful search API. The number of papers downloaded
per paper type category was as follows: 200 Essay, 99 Correspondence, 174
Synopsis, 200 Perspective, 74 Viewpoint, 93 Opinion, 107, Case Study, 312
Review, 200 Research.  These formed a corpus of 1459 papers.

@@

@@

Machine learning to build a classifer of paper type was conducted using random
forest learning \cite{Breiman2001}.  A random forest learning algorithm was
chosen because this is a fast, accurate and widely accepted learning
technology, and because the underlying decision trees can be used to describe
and inspect some of the reasoning behind the predictions made by the final
paper type classifiers.  The feature set consisted of the percentage
composition of the CoreSC labels given to the sentences in each paper (the
percentage of {\em Background} sentences, the percentage of {\em Hypothesis}
sentences, etc.).  Thus there are XX features corresponding to the XX CoreSC
labels.  The random forest learning was conducted using the Orange data mining
library for Python \cite{Curk2005}.  The parameters for the random forest
learner were as the defaults, except with min\_subsets=5 and same\_majority\_pruning=true.  We used 10-fold cross validation to estimate the precision,
recall and F-measure.

>>>>

\usepackage{subcaption}
\usepackage{soul}
\captionsetup{compatibility=false}

@@

\section{Methods}
\label{sec:2}
%how types were chosen,

\subsection*{Collection of scientific articles}
Partridge allows users to upload any paper which is free of copyright restrictions.
For the purpose of this study we needed a large set of papers that we could label with CoreSC
to investigate how this information assists classification into paper type.
Open Access (OA) journals \st{are research journals that provide their papers free
to read online, often making revenue by charging authors to publish their
articles \cite{solomon2012study}.  Many OA journals} provide free read access to
their papers, but do not permit the user of articles for data mining purposes.
The PLoS journals\footnote{PLoS: \url{http://www.plos.org/}} contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the Pubmed Central markup schema, which is compatible with SAPIENTA,
for papers published through their journal.  The PLoS journals advanced search
offers approximately 50 types of paper through which to restrict the search.
Many of these paper type categories contain too few papers to be useful, others
are too specific (e.g.``Message from PLoS'') or appear rather adhoc. 
Indeed it is not clear how these paper types have been identified. % in conclusions add how this can give recommendations for Plos and other journals
We chose to look at a range of types, some of which
we expect to overlap or have an unclear distinction.
These are namely: Essay, Correspondence, Case Study, Synopsis, Perspective, Viewpoint, Opinion,
Review, Research). 
%how papers were acquired 
The papers were downloaded from PLoS using {\em plosget.py}, a short python
script that uses the PLoS restful search API. The number of papers downloaded
per paper type category was as follows: 200 Essay, 99 Correspondence, 174
Synopsis, 200 Perspective, 74 Viewpoint, 93 Opinion, 107 Case Study, 312
Review, 200 Research.  These formed a corpus of 1459 papers.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX \cite{PDFX}, a free service hosted
by the University of Manchester (\url{http://pdfx.cs.man.ac.uk/}).
PDFX also provides output that already has some metadata, such as
title, author, and abstract, associated with it.  SAPIENTA allocates
each sentence within a paper a separate CoreSC label.  However, the
sentence boundaries must be detected before the document is passed
into SAPIENTA for annotation.  The sentence splitter uses a regular
expression to split the sentences within the document and then adds
the necessary \verb|<s>| tags to the markup to indicate the location
of each sentence.  Once the paper has been split, SAPIENTA is used to
annotate each sentence with a relevant CoreSC label.  These labels are
assigned using a conditional random fields model based upon a number
of features such as the sentence's location within the paper and pairs
and triplets of words found consecutively within the sentence
\cite{Liakata2012}.  The SAPIENTA system has been trained on a corpus
of 265 chemistry and biochemistry papers and we have done no domain
adaptation for the papers in PLoS.

@@

Machine learning to build a classifer of paper type was conducted using random
forest learning \cite{Breiman2001}.  A random forest learning algorithm was
chosen because this is a fast, accurate and widely accepted learning
technology, and because the underlying decision trees can be used to describe
and inspect some of the reasoning behind the predictions made by the final
paper type classifiers.  The feature set consisted of the percentage
composition of the CoreSC labels given to the sentences in each paper i.e. (the
percentage of {\em Background} sentences, the percentage of {\em Hypothesis}
sentences, etc.).  Thus there are 11 features corresponding to the 11 CoreSC
labels.  The random forest learning was conducted using the Orange data mining
library for Python \cite{Curk2005}.  The parameters for the random forest
learner were as the defaults, except with min\_subsets=5 and same\_majority\_pruning=true.  
We used 10-fold cross validation to estimate the precision,
recall and F-measure.
